Anxiety: a decision-theoretic perspective

### 2019-02-03 manuscript/manuscript.tex
rewrote results 3.1

environment\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript, palatino}
\usepackage{graphicx}
\usepackage{amsfonts, amsmath}
\usepackage{algorithm, algpseudocode}%

@@

In the following sections, we present the results of simulations of pessimistic
learning in a variety of MDP environments. In all simulations, the state-action
values under the optimal policy, $Q_*(s,a)$, were solved for directly through
Q-value iteration \citep{SuttonBarto1998,SuttonBarto1998,bertsekas2005}. The
details of this algorithm are presented in the appendix. Model parameters are
reported for each environment in their corresponding sections. All simulations
were carried out using the scientific python distribution and are publicly available
at on Github (https://github.com/szorowi1/SecretFunTimes).

@@

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
  }
  \caption{\textbf{Open Field}}
  \par Current parameters: Reward = 10, Shock = -10, gamma = 0.95.
\end{figure}

@@

The long-run state values under different beliefs are plotted in Figure 1. As
expected, under optimistic beliefs (Figure 1b) the entire environment sans negative
state takes on a positive value (close to the positive state itself). As this belief
breaks down, we observe a different pattern in the long-run state values. When
future occupancy of the negative state becomes a non-zero probability, the undesired
action transition exerts influence on action-outcome calculations, back-propagating,
and thereby infecting antecedent states. As the agent's belief becomes increasingly
pessimistic, the more the worst possible action exerts influence and the more
negative, on average, the state-space grows.

In the breakdown of perceived control then, negative states can grow worse (i.e.
more predictive of future negative outcome) and otherwise positive states can too.
This occurs as a simple result that if control not guaranteed, cannot guaranteed
bad things wont happen.

% I think you want to explain second order conditioning in PTSD in this context.

@@

\section{Appendix}
\subsection{Algorithms}

@@

\begin{algorithm}
  \caption{Betamax Temporal Difference Learning}

  \State Algorithm parameters: step size $\eta \in (0, 1)$, small $\epsilon > 0$
  \State Initialize $Q(s,a)$, for all $s \in S$, $a \in A$ arbitrarily, except that $Q(terminal,\cdot) = 0$
  \State
  \Loop \ for each episode
    \State Initialize $S$
    \While{$S \notin S(terminal)$}
      \State Choose $A$ from $S$ using policy derived from $Q$
      \State Take action $A$, observe $R, S$
      \State $Q(S, A) \leftarrow Q(S, A) + \eta \left[ R + \gamma \max_a Q(S',a) − Q(S, A) \right] $
      \State $S \leftarrow S'$
    \EndWhile
  \EndLoop

>>>>

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript, palatino}
\usepackage{graphicx}
\usepackage{amsfonts, amsmath}
\usepackage{algorithm, algpseudocode}%

@@

In the following sections, we present the results of simulations of pessimistic
learning in a variety of MDP environments. In all simulations, the state-action
values under the optimal policy, $Q_*(s,a)$, were solved for directly through
Q-value iteration \citep{SuttonBarto1998,SuttonBarto1998,bertsekas2005}. The
details of this algorithm are presented in the appendix. Model parameters are
reported for each environment in their corresponding sections. All simulations
were carried out with the python programming language and are publicly available
on Github (https://github.com/szorowi1/SecretFunTimes).

@@

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/01_field.png}}%
  }
  \par Figure 1: The open field environment is a large deterministic gridworld sans one rewarding
  (blue) and aversive (red) state (A). Under optimistic learning conditions (B), all states except
  for the harmful one become positive. Under pessimistic conditions (C), negative value spreads
  from the source to its antecedent states. The extent of the spread is worse for increased
  pessimism (D). (Parameters: $\gamma = 0.95$)
\end{figure}

@@

The same cannot be said for a the pessimistic agent. If an agent doubts its future
ability to act in according with the optimal policy, then it cannot rule out that,
in the future, it may inadvertently encounter the harmful state. This is more so
likely if the agent is already proximal to the harmful state. Under this pessimistic
belief, the possibility of future harm is no longer zero, as it is for the optimistic
agent, and thus negative value can propagate from the states antecedent to harm,
back to their respective antecedents and so on. The extent of this spread will depend
on the strength of the pessimistic belief (Figures 1c, 1d). With increasing pessimism,
harmful outcomes are assigned greater credence and thus exert greater and wider
influence on antecedent states.

In sequential decision environments then, a perceived lack of self-efficacy can
result in the propagation of negative value from potential future harms. Such a
pessimistic agent may correspondingly report an increased likelihood of threat
(i.e. a greater number of threatening states), increased severity of threat (i.e.
states with more negative value), and fear of states antecedent to the fear source.
As such, pessimistic learning can account for two of the hallmark symptoms of
the anxiety disorders.

@@

\section{Appendix}

@@
